---
title: "STA457 Assignment"
author: 'Isaac Aktam, Student ID: 998986575'
date: "August 05, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r dataImport, echo=FALSE, fig.width= 8, fig.height=4, fig.align='center'}
# use setwd to set your working directory and put the gasfur.csv there
 dat<-read.csv("gasfur.csv", header=TRUE)
 gas<-ts(dat[,1],freq=1, start=1)
 co2<-ts(dat[,2], freq=1, start=1)
# head(dat)
# par(mar=c(3,3,1,1))
# plot(cbind(co2,gas), main="")
```

#### Model
Fit the following transfer function noise model
$$ co2_t =\alpha + \sum^s_{j} v_j \cdot gas_{t-j-b} + e_t,     (1)$$
where $e_t$ is serially correlated.

#### 1. Use the ideas of prewhitening taught in class to identify b and s. 
According the following figure, I identify that $b=?$ and $s=?$.
```{r}
# Type R codes used to produce the figure for identifying b and s
install.packages("markdown", repos='http://cran.utstat.utoronto.ca/')
# library(markdown)
require(markdown)
install.packages("forecast", repos='http://cran.utstat.utoronto.ca/')
library(forecast)
ccf(gas, co2)
acf(gas)
pacf(gas)
acf(co2)
pacf(co2)

ar5model = arima(gas, order = c(5,0,0), include.mean = FALSE)
ar5model # Prints the AR(5) coefficients for gas
f_co2 = c(1,-ar5model$coef[1:5]) #create a filter to transform co2
f_co2
co2_filtered = filter(co2, f_co2, method=c("convolution"),sides=1)
co2_filtered
co2_prew = co2_filtered[6:296] # Transformed co2
co2_prew
gas_prew = ar5model$residuals[6:296] # Transformed gas
gas_prew
acf(gas_prew)
pacf(gas_prew)
CCF = ccf(co2_prew, gas_prew) #retrieve cross-correlations
CCF
vk = (sd(co2_prew)/sd(gas_prew))*CCF$acf # impulse response function
vk
ACF = acf(co2_prew)
plot(CCF, ylab = "CCF",main = "Cross-correlations after prewhitening")

#According to CCF plot, I identify that s = 4, b = 3
```

#### 2. Fit Eqn. (1) based on your preliminary identification and R arima function
Type the mathematical form of your model here
# We choose ARIMA(3,1,2) since it has the lowest aic. Also, its ccf shows that all of its values are within 95% CI. Therefore, its error terms behave like white noise.
# Fit for the error terms
$$  (1-03796*B-0.6830*B^2+0.5802*B^3)*(1-B)*e_{t} = (1+0.2184*B-0.6421*B^2)*j_{t}  $$
$$  (1-1.3796*B-0.3034*B^2+1.2632*B^3-0.5802*B^4)e_{t} = j_{t}+0.2184*j_{t-1}-0.6421*j_{t-2}  $$
$$  e_t = 1.3796*e_{t-1}+0.3034*e_{t-2}-1.2632*e_{t-3}+0.5802*e_{t-4}+j_{t}+0.2184*j_{t-1}-0.6421*j_{t-2}        $$
# Therefore, our final equation is
$$  co2_t = -.5249*gas_{t-3}-0.6202*gas_{t-4}-0.8556*gas_{t-5}-0.4746*gas_{t-6}-0.3414*gas_{t-7}+1.3796*e_{t-1}+0.3034*e_{t-2}-1.2632*e_{t-3}+0.5802*e_{t-4}+j_{t}+0.2184*j_{t-1}-0.6421*j_{t-2} $$
where $j_{t}$ ~ N(0,0.05645)


```{r}
# Lagging the gas

gas_lag3 = c(NA, NA, NA, gas[1:293])
gas_lag4 = c(NA, NA, NA,NA, gas[1:292])
gas_lag5 = c(NA, NA, NA,NA,NA, gas[1:291])
gas_lag6 = c(NA, NA, NA,NA,NA,NA, gas[1:290])
gas_lag7 = c(NA, NA, NA,NA,NA,NA,NA, gas[1:289])
gas_lag_master = cbind(gas_lag3,gas_lag4,gas_lag5,gas_lag6,gas_lag7)
colnames(gas_lag_master) = paste("gas_lag",3:7, sep = "")

gas_lag_master


# Need to choose optimal model with lowest abs(AIC)

# Run for loop to find the model with optimal p,d,q

aic_c = c()
p_c = c()
d_c = c()
q_c = c()
for (p in 1:5){
  for (d in 0:1){
    for (q in 1:5){
      p_c = c(p_c, p)
      d_c = c(d_c, d)
      q_c = c(q_c, q)
      fit_reg = arima(co2, xreg = gas_lag_master, order = c(p,d,q))
      print(fit_reg)
      print(c(p,d,q))
      aic_c = c(aic_c, fit_reg$aic)
    }
  }
}
aic_c
master_data = cbind(p_c,d_c,q_c,aic_c)
which.min(master_data[,4]) #row location of min aic
master_data[27,] # Get the row with the p, d, and q values related to min aic

fit_0 = arima(co2, xreg = gas_lag_master, order = c(3,1,2))
fit_0 #aic is 11.34
pacf(fit_0$residuals, na.action = na.pass)

fit_1 = auto.arima(co2, xreg = gas_lag_master)
fit_1 #aic is 14.4
pacf(fit_1$residuals, na.action = na.pass)


fit_2 = arima(co2, xreg = gas_lag_master, order = c(4,0,0))
fit_2 # aic is 14.68
pacf(fit_2$residuals, na.action = na.pass)

fit_3 = arima(co2, xreg = gas_lag_master, order = c(7,1,4), include.mean = FALSE) # testing of various p,d,q
fit_3 # aic is 11.55
pacf(fit_3$residuals, na.action = na.pass) # Seems to have best PACF graph so far

aic_list = list(ARIMA312 = fit_0$aic, ARIMA410 = fit_1$aic, ARIMA400 = fit_2$aic, ARIMA714 = fit_3$aic)
aic_list

# Choose model fit_0 since it has the lowest aic of 11.34


# Show R codes that estimate your model and the corresponding estimation result
```

#### 3. Checking model adequacy of your fitted model.
The analysis of the model adequacy of the fitted model is as follows:

* e.g. My model pass the residual correlation check using ?? because the corresponding p-value is ??
* e.g. The cross correlation between xx and xx .....

```{r}
# Show your R codes and results for checking model adequacy

install.packages("TSA", repos='http://cran.utstat.utoronto.ca/')
library("TSA")
p_value_list = list(ARIMA_3_1_2 = LB.test(fit_0, type = c("Ljung-Box"))$p.value, ARIMA_4_1_0 = LB.test(fit_1, type = c("Ljung-Box"))$p.value, ARIMA_4_0_0 = LB.test(fit_2, type = c("Ljung-Box"))$p.value,  ARIMA_7_1_4 = LB.test(fit_3, type = c("Ljung-Box"))$p.value)
p_value_list
# All of my models pass the residual correlation check using Ljung-Box test because all of the corresponding p-values are greater than 0.05
# The cross correlation between gas and residuals of ARIMA(3,1,2) model shows that all of the values are within 
# the 95% CI, or upper and lower threshold 

ccf(gas, fit_0$residuals, na.action = na.pass)
ccf(gas, fit_1$residuals, na.action = na.pass)
ccf(gas, fit_2$residuals, na.action = na.pass)
ccf(gas, fit_3$residuals, na.action = na.pass)

# From the graph of ccf's, we can see that fit_2 (ARIMA(4,0,0)) has the least optiomal CCF graph.Therefore, we need to choose the most optimal model from: 
# fit_0, ARIMA(3,1,2)
# fit_1, ARIMA(4,1,0)
# fit_3, ARIMA(7,1,4)
# From the above 3 models, fit_0, ARIMA(3,1,2), has the least aic, therefore
# it is reasonable to choose it as the most optimal model
# Also, the ccf(gas, fit_0$residuals) graph shows that residuals behave like white noise and thus have no spikes outside of 95% CI. Beautiful graph indeed.


```

